#Sampling, Probability distributions and bootstrapping {#lab6bsampling_bootstrapping}

```{r setup18, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r}
library(tidyverse)
```

## Data 
```{r}
data_students <- read.csv("https://userpage.fu-berlin.de/soga/200/2010_data_sets/students.csv")
```

## Probability distributions and predictions

We first calculate the sample mean ($\overline{x}$), and the standard deviation of the sample mean $S_x$, of the target variable (in this case the `height` variable in the `data_students` dataframe. Then, we standardize the `height` variable to get a standard normal distribution with =0 and =1 and assign it to an appropriate variable name.

The formula for culculating the standard deviation of the sample mean is:

$s_x = \sqrt{\frac{\sum_{i=1}^N (x_i - \overline{x})^2}{N-1}}$

## EXERCISE 1. Calculating the stardard deviation

 a) Calculate with R the standard deviation of the sample mean by writing an expression in R that implements the above formula.

 b) Use the function `mean()` calculate the sample mean, store it in a variable in R
 
 c) and the function `sd()` to confirm your 'manual' calculations

```{r}
## a)
height_values <- data_students$height
degrees_freedom <- length(height_values)-1
sum_of_squares <- (height_values - mean(height_values))^2 %>% sum() 
sd = sqrt(sum_of_squares / degrees_freedom) 
sd

## b
mean_height <- mean(height_values)

## c
sd_height <- sd(height_values)

```

--- END OF EXERCISE ---

## Z-transformation of the female observations
Here we transform the female sample distribution to a standard normal distribution.
```{r}
females <- data_students %>%
  dplyr::filter(gender == "Female") %>%
  as_tibble()

females

mean_height_f <- mean(females$height)
sd_height_f <- sd(females$height)

height_z <- (females$height - mean_height_f)/sd_height_f

height_z %>% hist(main = "Z-transformed, female height")

```

The female height variable has a mean of `r mean(females$height)` cm of and a standard deviation of `r sd(females$height)` cm.

Finding the area to the left of a specified z
-score or x
-score

Question 1
What is the probability of a randomly picked female student from the student data set with a height less or equal to 168 cm. Thus, we are looking for P(x≤168)

First, we calculate the probability for the standardized variable. Therefore, we have to transform our value of interest (168 cm) into a z-score.

$z = \frac{x−μ}{σ} = \frac{168−163.7}{7.9} = 0.55$

Then we have to calculate the area under the curve left to the obtained z

the area under the curve of a normal distributed variable can be calculated by applying the pnorm function. The pnorm function is written as pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE). For this particular example we can accept all default argument values.

```{R}
x <- 168 # height in cm 
x_z <- (x - mean_height_f)/sd_height_f # z-transformation
pnorm(x_z) %>% round(2)
```

## [1] 0.7084447

Awesome, we have a result: P(z≤0.55)≈0.71

Now, we do the same calculation, however this time we skip the step of standardization. Thanks to the power of R we do not need to rely on tables, but can we can easily put the sample mean, $\overline{x}$ and the sample standard deviation, $s_x$ into the pnorm function.

```{r}
x <- 168 # height in cm 
pnorm(x, 
      mean = mean_height_f, 
      sd = sd_height_f, 
      lower.tail = TRUE, log.p = FALSE
      ) %>%
  round(3)

## [1] 0.7084447
```

Perfect! The numbers match: P(x≤168)≈0.71. To make sure we realize what is going on, both the area under the curve for the standardized variable in z-values (left panel) as well as the area for the non-standardized variable in cm (right panel) are visualized below.

```{r}

ggplot(females, aes(x = height_z)) +
stat_function(
fun = dnorm,
args = with(females, c(mean = mean(height_z), sd = sd(height_z)))
) + stat_function(
  args = with(females, c(mean = mean(height_z), sd = sd(height_z))),
    fun = dnorm,
    geom = "area",
    fill = "steelblue",
    alpha = .3,
  xlim = c(-4, 0.55)) +
  ggtitle("P(z =< 0.55")



ggplot(females, aes(x = height)) +
stat_function(
fun = dnorm,
args = with(females, c(mean = mean(height), sd = sd(height)))
) + stat_function(
  args = with(females, c(mean = mean(height), sd = sd(height))),
    fun = dnorm,
    geom = "area",
    fill = "steelblue",
    alpha = .3,
  xlim  = c(140, 168)) +
  ggtitle()

```


