---
title: "Course Activities"
author: "Marc A.T. Teunis"
date: "3/27/2019"
output:
  ioslides_presentation: 
    highlight: pygments
    widescreen: yes
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, root_1, include=FALSE}
## defines the root of the project for later use
require("rprojroot") || utils::install.packages("rprojroot")
library(rprojroot)
root <- find_root_file(criterion = is_rstudio_project)
```

```{r, include=FALSE}
exercise_files <- list.files(path = file.path(root, "exercises"), full.names = TRUE, recursive = TRUE)

presentation_files <- list.files(path = file.path(root, "presentations"), full.names = TRUE, recursive = TRUE)

##exercise_files
```

## Course Overview (this document)

 - The course overview provides a summary of the course. It is a document that is fluid throughout the course. It is the current document before you. I will make changes to this as we go along to personalize the course to your needs as much as possible.
  - The course workflow will be addressed
  - If applicable the Course Exam will be addressed (Bachelor Course Only)
  - We will get to know some of the conventions used in the course materials to aid readibility and completion of the exercises

## Menti.com

<div style='position: relative; padding-bottom: 56.25%; padding-top: 35px; height: 0; overflow: hidden;'><iframe allowfullscreen='true' allowtransparency='true' frameborder='0' height='315' src='https://www.mentimeter.com/embed/45fa38401d60a4886d2e21c9f2ebc240/d2a9c72548cf' style='position: absolute; top: 0; left: 0; width: 100%; height: 100%;' width='420'></iframe></div>


## Labs

 - Each day (lesson) consists of several labs
 - Is a combination of direct instruction and introduction to a new topic  
 - Each topic is accompanied by exercises. 
 - Some of the exercise will be addressed during instruction. 
 - Some exercises will have to be prepared outside of class. 
 - Each consecutive lesson will follow up on the previous exercises. 
 - During the follow up solution for the problems that were encountered during the preparation of the exercises will be provided 
 - **There is always room for questions!**

## Cases

 - During each lesson I will ask for a case from the participants
 - Some problem you are working on that is relevant for the topics in that day's class
 - If possible I will prepare a short demo, workflow or provide some tips on how to tackle the case problem
 - I will ask for your input on the matter
 
# Day 1

 1. Course overview, Datamanagement 
 2. Git, Github.com  
 3. Visualization examples in R RStudio
 4. RMarkdown
 5. R basics and Getting Help
 6. Intro to R 

## Lab 1 Course intro
  
  - We will discuss some of the conventions used in the slides and talk a little bit more about R and how to get additional resources for learning to use R
  - You will intiate your first R-package
  - To find some of the resources used in this course and for progessing you R learning: look at the "resources" file in the root. 
  - In RStudio, important tools for reproducibility of (data)-research are build-in: `{rmarkdown}`, RStudio Projects, the `{here}` package 
 - Every data-project has files, and often many, many files
 - We will see a short list of rules to keep track of things and to aid reproducibility and tracebility
 - You will learn about the structure of an R package to aid reproducibility

## Lab 2 Git/github.com

 - Git is used to keep track of changes, collaborate and reproduce.
 - Github is used to share code (any code), data and other files and find additional great R resources and packages  
 - Git/Github are is a widely used platform for the development of software

## Lab 3 Visualizations (demo)

 - Are a great way to get inspired on the possibilities for using R
 - Are great for exploring new datasets
 - Are a great way to learn (new) ways of writing R code

## Lab 4 RMarkdown

 - All Exercises and demos are RMarkdown files
 - We will study this format 
 - During this lab you will learn (some of) the possibilities of RMarkdown
 - You will create your own RMarkdown and push it to you Github repository

## Lab 5 RStudio

 - We will use RStudio Server as an Integrated Development Environment (IDE) for developing in R
 - RStudio is one of most widely used IDEs
 - RStudio integrates with the `Tidyverse`
 - RStudio enables easy creation of your own R package
 
 
## Lab 6 Intro to R

 - You will learn the basics about R-commands (called `functions` in R)
 - We will see the most common objects (Vectors, Lists, Dataframe, Matrix)
 - You will learn how the help function for functions and packages works

# Day 2  

## Activities

 - Intro, file management, course conventions
 - Continue lab 6
 - Start with Data wrangling - lab 7
 - Start with Visualizations - lab 8
 






