---
title: "Data Management - Demo"
author: "Marc Teunis"
date: "`r Sys.Date()`"
output:
  ioslides_presentation: 
    highlight: pygments
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Files

 - When doing data analysis you will gather A LOT OF FILES.
 - Having no system will cause you time and will lead to loss of data and information

## File names
 
 - Use "Linux" style file names: "snake_case.txt"
 
## A few simple rules

 - Space is cheap, confusion is expensive
 - Do everything via code
 - Write code that runs from begin to end
 - Never change 'raw'-data or filename
 - Set guidelines with the group and stick to them
 
  __after Edna Ridge - "Guerilla Analytics"__

## Project structure

Use a flat folder structure

```{r}
fs::dir_tree(here::here("participant_cases"))
```




## Naming Conventions

 - I use `snake_case` for variables, function names and file names
 - `camelCase` is also used
 - `dot.case` is not modern 
 - DO NOT rename files (raw-data & course files)
 
 
## Recieving Data


 